Probability

### 2016-04-30 prob.tex
Add citation, but need to verify the material...

In a strict computational or experimental setting, one may question
the reliance of the definition of probability space on the uncountable
and uncomputable real interval $[0,1]$. This interval includes numbers
like $0.h_1h_2h_3\ldots$ where $h_i$ is 1 or 0 depending on whether
Turing machine $M_i$ halts or not. Such numbers cannot be
computed. This interval also includes numbers like $\frac{\pi}{4}$
which can only be computed with increasingly large resources as the
precision increases.
% \yutsung{Check the meaning of the ``computing $\frac{\pi}{4}$'', because
% Bailey Borwein Plouffe formula can computing the $n$th binary digit of $\pi$ 
% using base 16 directly.}
Therefore, in a resource-aware setting, it is more appropriate to
consider probability measures that map events to a finite set of
elements computable with a fixed set of resources. We will consider
two approaches: set-valued probability
measures~\cite{Artstein1972,PuriRalescu1983} and interval-valued
probability measures~\cite{JamisonLodwick2004,Weichselberger2000}.

@@

In interval-valued probability measures, if $\mu\left(E\right)=\left[a,b\right]$,
the left-endpoint~$a$ called belief denotes the strength of the
evidence that $E$ happens, and the right-endpoint~$b$ called plausibility
denotes the strength of the evidence that contradicts $E$. If we
define $1-\left[a,b\right]=\left[1-b,1-a\right]$, then we intuitively
have 
\begin{eqnarray*}
\pmeas\left(\emptyset\right) & = & \imposs\\
\pmeas\left(\Omega\right) & = & \necess\\
\pmeas\left(\Omega\backslash E\right) & = & 1-\pmeas\left(E\right)
\end{eqnarray*}
Next, if we define $\sum_{i}\left[a_{i},b_{i}\right]=\left[\sum_{i}a_{i},\sum_{i}b_{i}\right]$,
then 
\begin{itemize}
\item for a collection $E_{i}$ of pairwise disjoint events, we have $\pmeas\left(\bigcup_{i}E_{i}\right)\subseteq\sum_{i}\pmeas\left(E_{i}\right)$. 
\end{itemize}
Notice that the equality may not hold in general This statement says
that the evidences of $\bigcup_{i}E_{i}$ is at least as strong as
putting all the evidences of $E_{i}$ together, but some evidence
may only be acquired for $\bigcup_{i}E_{i}$ as the whole. Therefore,
$\pmeas\left(\bigcup_{i}E_{i}\right)$ is a subset of $\sum_{i}\pmeas\left(E_{i}\right)$,
but may not equal. In our example, 
\begin{eqnarray*}
 &  & \pmeas(\{HH,TH\})=\necess=[1,1]\\
 & \subseteq & \left[\frac{1}{2},\frac{3}{2}\right]=\left[0,\frac{1}{2}\right]+\left[\frac{1}{2},1\right]=\unlikely+\likely=\pmeas(\{HH\})+\pmeas(\{TH\})\textrm{ .}
\end{eqnarray*}
However, $\pmeas(\{HH,TH\})$ is a proper subset of $\pmeas(\{HH\})+\pmeas(\{TH\})$
because if we check the complement of $\{HH,TH\}$, we have 
\begin{eqnarray*}
 &  & \pmeas(\{HH,TH\})=1-\pmeas(\{HT,TT\})=1-\imposs=\necess\textrm{ ,}
\end{eqnarray*}
and $\pmeas(\{HT,TT\})=\imposs$ cannot be used to reasoning the probability
of $\{HH\}$ and $\{TH\}$ individually.

\yutsung{Although this is clearer than the previous one, I keep some
intervals instead write everything in \imposs, \unlikely, \likely,
and \necess. The biggest point is the addition rule~$\sum_{i}\left[a_{i},b_{i}\right]=\left[\sum_{i}a_{i},\sum_{i}b_{i}\right]$:
\begin{itemize}
\item This rule without \overflow will be push complexity to ``we may abort
computing probability when....''
\item It seems more concisely than a table among \imposs, \unlikely, \likely,
and \necess.
\end{itemize}
However, if you prefer as less interval as possible, I may try to
replace them by \imposs, \unlikely, \likely, and \necess...}
\amr{I like this. We just might need to rewrite a little but overall
  it's good.}

@@

\amr{give example non-commuting events}

\yutsung{Isn't $\proj{0}$ and $\proj{\ps}$ an example in the example of
One-qubit quantum probability space.}

@@

\bibliographystyle{plain}
\bibliography{discreteGBKS}

>>>>

In a strict computational or experimental setting, one may question
the reliance of the definition of probability space on the uncountable
and uncomputable real interval $[0,1]$. This interval includes numbers
like $0.h_1h_2h_3\ldots$ where $h_i$ is 1 or 0 depending on whether
Turing machine $M_i$ halts or not. Such numbers cannot be
computed. This interval also includes numbers like $\frac{\pi}{4}$
which can only be computed with increasingly large resources as the
precision increases.
% \yutsung{Check the meaning of the ``computing $\frac{\pi}{4}$'', because
% Bailey Borwein Plouffe formula can computing the $n$th binary digit of $\pi$ 
% using base 16 directly.}
Therefore, in a resource-aware setting, it is more appropriate to
consider probability measures that map events to a finite set of
elements computable with a fixed set of resources. We will consider
two approaches: set-valued probability
measures~\cite{Artstein1972,PuriRalescu1983} and interval-valued
probability measures~\cite{Dempster1967,Weichselberger2000,JamisonLodwick2004}.

@@

In interval-valued probability measures, if $\mu\left(E\right)=\left[a,b\right]$,
the left-endpoint~$a$ called belief denotes the strength of the
evidence that $E$ happens, and the right-endpoint~$b$ called plausibility
denotes the strength of the evidence that contradicts $E$. If we
define $1-\left[a,b\right]=\left[1-b,1-a\right]$, then we intuitively
have 
\begin{eqnarray*}
\pmeas\left(\emptyset\right) & = & \imposs\\
\pmeas\left(\Omega\right) & = & \necess\\
\pmeas\left(\Omega\backslash E\right) & = & 1-\pmeas\left(E\right)
\end{eqnarray*}
Next, if we define $\sum_{i}\left[a_{i},b_{i}\right]=\left[\sum_{i}a_{i},\sum_{i}b_{i}\right]$,
then 
\begin{itemize}
\item for a collection $E_{i}$ of pairwise disjoint events, we have $\pmeas\left(\bigcup_{i}E_{i}\right)\subseteq\sum_{i}\pmeas\left(E_{i}\right)$. 
\end{itemize}
Notice that the equality may not hold in general. This statement says
that the evidences of $\bigcup_{i}E_{i}$ is at least as strong as
putting all the evidences of $E_{i}$ together, but some evidence
may only be acquired for $\bigcup_{i}E_{i}$ as the whole. Therefore,
$\pmeas\left(\bigcup_{i}E_{i}\right)$ is a subset of $\sum_{i}\pmeas\left(E_{i}\right)$,
but may not equal. In our example, 
\begin{eqnarray*}
 &  & \pmeas(\{HH,TH\})=\necess=[1,1]\\
 & \subseteq & \left[\frac{1}{2},\frac{3}{2}\right]=\left[0,\frac{1}{2}\right]+\left[\frac{1}{2},1\right]=\unlikely+\likely=\pmeas(\{HH\})+\pmeas(\{TH\})\textrm{ .}
\end{eqnarray*}
However, $\pmeas(\{HH,TH\})$ is a proper subset of $\pmeas(\{HH\})+\pmeas(\{TH\})$
because if we check the complement of $\{HH,TH\}$, we have 
\begin{eqnarray*}
 &  & \pmeas(\{HH,TH\})=1-\pmeas(\{HT,TT\})=1-\imposs=\necess\textrm{ ,}
\end{eqnarray*}
and $\pmeas(\{HT,TT\})=\imposs$ cannot be used to reasoning the probability
of $\{HH\}$ and $\{TH\}$ individually.

@@

@@

\bibliographystyle{unsrt}
\bibliography{discreteGBKS}
